A Temporal Logic Based Framework for Intrusion Detection

### 2023-09-09 Paper/main.tex
update

\subsection{Relationship to LTL+P}
According to~\cite{barringer2004rule,barringer2003eagle}, we can define both \textbf{Future Time LTL} and \textbf{Past Time LTL} operators. The future modalities are written in section~\ref{sec:ltl}, therefore let us list the past ones:
\begin{align*}
    & \underline{\text{min}}\ \mathtt{Previous}(\underline{\text{Form}}\ F) = \odot F \\
    & \underline{\text{max}}\ \mathtt{AlwaysInPast}(\underline{\text{Form}}\ F) = F \land \odot \mathtt{AlwaysInPast}(F) \\
    & \underline{\text{min}}\ \mathtt{EventuallyInPast}(\underline{\text{Form}}\ F) = F \lor \odot \mathtt{EventuallyInPast}(F) \\
    & \underline{\text{min}}\ \mathtt{Since}(\underline{\text{Form}}\ F_1, \underline{\text{Form}}\ F_2) = F_2 \lor (F_1 \land \odot \mathtt{Since}(F_1,F_2))
\end{align*}
% Note that UNLESS is maximal
By combining the definitions for the future and past time LTLs defined above, we obtain a temporal logic over the future, present and past, in which one can freely intermix the future and past time modalities; thus, we can state that EAGLE is expressive at least as LTL+P.
% a differenza di ltl non salva tutta la sequenza ma solo una formula che riassume tutto. La seconda componente non è uno stato (insieme di cose vere) ma un record
% EAGLE è almeno espressivo quanto LTLP

@@

\textbf{Cookie-stealing attacks.}
A cookie is a technique for tracking sessions that a web application server sends to a web client and uses to keep track of client-specific session data. These cookies that serve as authentication tokens are automatically included in requests by clients and in this scenario we assume that a session is exclusively identified by the client's IP address. When a malicious user uses an old cookie provided to a different IP address to take over a session, it is considered an attack.In order to monitor this attack we need to look at a web-server (application-level) log that contains a record of all sessions that the server participates in, along with session-specific state information.
According to the formula below, a specific cookie must always be utilized by the same IP address.
%%  esempio sbagliato IMO
%\begin{align*}
%    & \underline{\text{min}}\ \mathtt{SafeUse}(\underline{\text{string}}\ c,\underline{\text{int}}\ i) = ((name=c)\rightarrow(ip=i))\land\odot\ \mathtt{SafeUse}(c,i) \\
%    & \underline{\text{mon}}\ \mathtt{CookieSafe} =\ \mathtt{Always}(\mathtt{SafeUse}(name,ip)) 
%\end{align*}
\begin{align*}
    & \underline{\text{min}}\ \mathtt{Hijack}(\underline{\text{string}}\ c,\underline{\text{string}}\ i) = ((name=c)\land\neg(ip=i))\lor\odot\ \mathtt{Hijack}(c,i) \\
    & \underline{\text{mon}}\ \mathtt{CookieSafe} =\ \mathtt{Always}(\neg\mathtt{Hijack}(name,ip)) 
\end{align*}
A trace that violates this formula therefore encodes a cookie-stealing attack. The rule $\mathtt{Hijack}$ checks wheter it occurs a situation in which we have the same coookie $c$ associated with two different ip $i$; if it's true then it means that a cookie has been stolen and reused by an attacker. With $\mathtt{CookieSafe}$ we want to check that it never occurs a situation just as the one described.

>>>>

\subsection{Relationship to LTL+P}
According to~\cite{barringer2004rule,barringer2003eagle}, we can define both \textbf{Future Time LTL} and \textbf{Past Time LTL} operators. The future modalities are written in section~\ref{sec:ltl}, therefore let us list the past ones:
\begin{align*}
    & \underline{\text{min}}\ \mathtt{Previous}(\underline{\text{Form}}\ F) = \odot F \\
    & \underline{\text{max}}\ \mathtt{AlwaysInPast}(\underline{\text{Form}}\ F) = F \land \odot \mathtt{AlwaysInPast}(F) \\
    & \underline{\text{min}}\ \mathtt{EventuallyInPast}(\underline{\text{Form}}\ F) = F \lor \odot \mathtt{EventuallyInPast}(F) \\
    & \underline{\text{min}}\ \mathtt{Since}(\underline{\text{Form}}\ F_1, \underline{\text{Form}}\ F_2) = F_2 \lor (F_1 \land \odot \mathtt{Since}(F_1,F_2))
\end{align*}
By combining the definitions for the future and past time LTLs defined above, we obtain a temporal logic over the future, present and past, in which one can freely intermix the future and past time modalities; thus, we can state that EAGLE is expressive at least as LTL+P.
% a differenza di ltl non salva tutta la sequenza ma solo una formula che riassume tutto. La seconda componente non è uno stato (insieme di cose vere) ma un record

@@

\textbf{Cookie-stealing attacks.}
A cookie is a technique for tracking sessions that a web application server sends to a web client and uses to keep track of client-specific session data. These cookies that serve as authentication tokens are automatically included in requests by clients and in this scenario we assume that a session is exclusively identified by the client's IP address. When a malicious user uses an old cookie provided to a different IP address to take over a session, it is considered an attack.In order to monitor this attack we need to look at a web-server (application-level) log that contains a record of all sessions that the server participates in, along with session-specific state information.
According to the formula below, a specific cookie must always be utilized by the same IP address.

\begin{align*}
    & \underline{\text{min}}\ \mathtt{Hijack}(\underline{\text{string}}\ c,\underline{\text{string}}\ i) = ((name=c)\land\neg(ip=i))\lor\odot\ \mathtt{Hijack}(c,i) \\
    & \underline{\text{mon}}\ \mathtt{CookieSafe} =\ \mathtt{Always}(\neg\mathtt{Hijack}(name,ip)) 
\end{align*}
A trace that violates this formula therefore encodes a cookie-stealing attack. The rule $\mathtt{Hijack}$ checks wheter it occurs a situation in which we have the same coookie $c$ associated with two different ip $i$; if it's true then it means that a cookie has been stolen and reused by an attacker. With $\mathtt{CookieSafe}$ we want to check that it never occurs a situation just as the one described.
